### 2023-03-03 v/401e10fd5dafceb54fd6b33a37b7c721ffce1eb8/images/104016/oj_scalebar.tex
Remove wrongly/forcibly added files [ci skip]
This build is based on
https://github.com/habi/EAWAG-manuscript/commit/401e10fd5dafceb54fd6b33a37b7c721ffce1eb8.

This commit was created by the following CI build and job:
https://github.com/habi/EAWAG-manuscript/commit/401e10fd5dafceb54fd6b33a37b7c721ffce1eb8/checks
https://github.com/habi/EAWAG-manuscript/actions/runs/4323430327

\documentclass[tikz]{standalone}
\usepackage{tikz}			% for drawing everything
	\usetikzlibrary{spy}	% for zooming
\usepackage{siunitx}		% for nice SI units
\usepackage{shadowtext}	% for shadowed text on the scalebar
	\shadowoffset{1pt}	% ideally the same as on line 13...
	\shadowcolor{green}	% ideally the same as on line 13...
\newcommand{\imsize}{\linewidth}% default width of image
\newlength\imagewidth% needed for correct scalebar
\newlength\imagescale% needed for correct scalebar
\begin{document}%
%----------
\tikzset{shadowed/.style={preaction={transform canvas={shift={(1pt,-1pt)}},draw=green, thick}}} % shadowed drawing https://tex.stackexchange.com/a/185853/828
\pgfmathsetlength{\imagewidth}{\imsize}%
\pgfmathsetlength{\imagescale}{\imagewidth/1153}%
\def\x{713}% scalebar-x starting at golden ratio of image width of 1153px = 713
\def\y{734}% scalebar-y at 90% of image height of 815px = 734
\def\mag{4}% magnification of inset
\def\size{75}% size of inset
\begin{tikzpicture}[x=\imagescale,y=-\imagescale,spy using outlines={rectangle,magnification=\mag,size=\size,connect spies}]
	\begin{scope}
		\clip (0,0) rectangle (1153,815);
		%\clip (576.5,407.5) circle (407.5);
		\node[anchor=north west, inner sep=0pt, outer sep=0pt] at (0,0) {\includegraphics[width=\imagewidth]{oj}};
	\end{scope}
	%\spy [red] on (853,515) in node at (0,0) [anchor=north west];
	% 200.664px = 0.865mm -> 100px = 431.069um -> 115.991px = 500um, 23.198px = 100um
	\draw[|-|,blue,thick] (533,336) -- (533,536) node [sloped,midway,above,fill=white,semitransparent,text opacity=1] {\SI{0.865}{\milli\meter} (201px) TEMPORARY!};
	\draw[|-|,white,thick,shadowed] (\x,\y) -- (\x+115.991,\y) node [midway,above] {\shadowtext{\SI{500}{\micro\meter}}};
	%\draw[color=red, anchor=south west] (0,815) node [fill=white, semitransparent] {Legend} node {Legend};
\end{tikzpicture}%
%----------
\end{document}%