### 2023-07-03 content/submission/submission/CoverLetter.Cichlids.PLOSONE.tex
Add cover letter

\documentclass[color,english,personal]{ubletter}

\begin{document} 
\begin{letter}{}
\subject{Submission of manuscript for publication}
\opening{To whom it may concern}

Herewith we submit our manuscript on the \emph{Microtomographic investigation of a large
corpus of cichlids} as an original contribution to PLOS ONE.

The manuscript presents how we efficiently and non-destructively acquired three-dimensional tomographic data from a large collection of cichlids from Lake Victoria in Africa.
We imaged the oral and pharyngeal jaws and the whole skull of these fishes as a preparation work for in-depth morphological description.
In the manuscript we describe our method and an present an outlook on analyzing the acquired data; a morphological description of the oral and pharyngeal jaws of the fishes, a three-dimensional geometric morphometrics analysis of landmark features on the skulls, and a robust method to automatically extract the otoliths of the fishes from the tomographic data.

A preprint of the manuscript is available on biorR\(\chi\)iv with the \href{https://doi.org/10.1101/2023.03.30.534917}{DOI:10.1101/2023.03.30.534917}.

We are looking forward to your positive feedback.

\addtolength{\medskipamount}{-\medskipamount} % skip some space with imported image
\closing{Yours sincerely\\\includegraphics[scale=0.618]{_Signature}}
\end{letter}
\end{document}